---
title: "615 Final Project"
subtitle: "Video Game Sales"
author: "Xinyi Wang"
date: "12/9/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Introduction

Video gaming has a long history as far back as the early 1950s but it did not reach mainstream popularity until the 1970s and 1980s. Video games are a billion-dollar business and have been for many years. In 2016, the video game market in the United States was valued at 17.68 billion U.S. dollars. 

The purpose of this project is to demonstrate what I have learned and my ability to extend knowledge as apply R in new situations - which is explore the development of the video game industry from 1983-2016.

#Read and Clean Data
```{r,warning=FALSE,message=FALSE,echo=FALSE}
library(dplyr)
library(ggplot2)
library(tidyverse)
library(tidyr)
setwd("/Users/CindyWang/Desktop/615/final_project/615final_project")
rawdata <- read.csv("video_game_sales.csv")
mydata <- rawdata

#Remove observations without "Name"
mydata$Name <- as.character(mydata$Name)
mydata <- filter(mydata, Name!="")

#Remove game released after 2017 and NA
mydata$Year_of_Release <- as.numeric(as.character(mydata$Year_of_Release))
mydata <- filter(mydata,Year_of_Release<2017,Year_of_Release>1982)
glimpse(mydata)
```

First step is to read and clean data. I removed all observations that do not have game names and released after 2016 since the dataset is from web scrape of VGChartz in late 2016.

#EDA

The next important part is data visualization. In this project, I will focus on publishers, game sales by region, platform, and genre. 

##1.Games Released Each Year by All Publishers
```{r,echo=FALSE}
mydata  %>%
  group_by(Year_of_Release)  %>%
  summarise(num_of_games = n_distinct(Name))  -> games_by_year

ggplot(games_by_year) +
  geom_point(mapping = aes(x=Year_of_Release, y=num_of_games),stat = 'identity') +
  geom_line(mapping = aes(x=Year_of_Release, y=num_of_games),stat = 'identity') + 
  labs(x = "Year", y = "Number of Games",title="Fig 1.Game Released Each Year by all Publishers")
```

We can see from Fig 1 that there are a lot of video games released during 2005-2008, which means the competition became intense in the late 2000s. 

##2.top 10 publishers with higher revenue 
```{r,echo=FALSE}
sales_per_publisher <- mydata %>%
  group_by(Publisher) %>%
  summarise(ttl_sales = sum(Global_Sales)) %>%
  arrange(desc(ttl_sales))
top10<- sales_per_publisher[1:10,]

ggplot(top10) + 
  geom_bar(aes(x = reorder(Publisher,ttl_sales),y = ttl_sales, fill = Publisher),stat = "identity")  +
  labs(x = "Publishers", y = "Total Global Sales(millions)",title="Fig 2.Top 10 Publishers with Higher Revenue ") + 
  theme(legend.position="none") +
  coord_flip() 
```

Fig 2 shows that form 1983 to 2016, Nintendo is the publisher who get highest revenue and follows by Electronic Arts and Activision. Surprisingly that Sony Computer Entertainment not reached half of Nintendo's total revenue, by looking into the whole publishers name I found that is more likely because Sony Company has a lot of branches such as "Sony Online Entertainment","Sony Music Entertainment",etc.

##3.Sales per region
```{r,warning=FALSE,message=FALSE,echo=FALSE}
aggr_coun <- mydata%>% 
  select(Year_of_Release, Genre, NA_Sales:Global_Sales) %>% 
  group_by(Year_of_Release, Genre) %>%
  summarise_each(funs(sum), NA_Sales:Global_Sales)

aggr_tab <- rbind(aggr_coun %>% mutate (Region = "JP", Sales = JP_Sales), 
                  aggr_coun %>% mutate (Region = "NA", Sales = NA_Sales), 
                  aggr_coun %>% mutate (Region = "EU", Sales = EU_Sales), 
                  aggr_coun %>% mutate (Region = "Other", Sales = Other_Sales) )

aggr_year <- aggr_tab %>% 
  select(Year_of_Release, Region, Sales) %>% 
  group_by(Year_of_Release, Region) %>% 
  summarise_each(funs(sum), Sales)

ggplot(data=aggr_year)+ 
  geom_line(aes(x=Year_of_Release,y=Sales,color=Region))+
  labs(y = "Sales", x = "Year of release",title="Fig 3.Sales per region")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
  
```

Fig 3 shows the sales trend for different regions from 1983 to 2016. It's clear that North America had the most drastic spike in the late 2000s.

##4.global sales map by platform and year
```{r,echo=FALSE,message=FALSE,warning=FALSE}
byPlatform <- mydata[,c("Year_of_Release","Platform","NA_Sales","EU_Sales","JP_Sales","Other_Sales")]
platform_count <- byPlatform %>%
  group_by(Year_of_Release,Platform) %>%
  summarise(NA_Sales=sum(NA_Sales),EU_Sales=sum(EU_Sales),JP_Sales=sum(JP_Sales),Other_Sales=sum(Other_Sales))
new <- gather(platform_count,area,sales,NA_Sales,EU_Sales,JP_Sales,Other_Sales)
new$Platform <- as.character(new$Platform)

##Take Platform="3DS" and Year_of_Release="2016" as example
ex_map_data <- new %>% 
  filter(Year_of_Release==2016,Platform=="3DS")

library(tmap)
# data(World)
wm <- map_data("world")
cc <- raster::ccodes()
mappings <- c("North America"="NA_Sales", "Europe"="EU_Sales") # you add the others here
cc$MYCONTINENTS <- mappings[cc$continent]
cc$MYCONTINENTS <- ifelse(cc$NAME=="Japan","JP_Sales",cc$MYCONTINENTS)
cc$MYCONTINENTS <- ifelse(is.na(cc$MYCONTINENTS),"Other_Sales",cc$MYCONTINENTS)
cc <- left_join(cc, ex_map_data, by = c("MYCONTINENTS"="area"))

## 31 country names need to be mapped... 
# wm$region %>% 
#   unique %>% 
#   setdiff(cc$NAME)
# ...                        
# [7] "Canary Islands"  "UK"  "Heard Island"     
# ...
## For example, UK is called United Kingdom in cc:
# unique(grep("Kingdom", cc$NAME, value=T, ignore.case=T))
# [1] "United Kingdom"
mappings <- c("UK"="United Kingdom", "USA"="United States","Ivory Coast"="Côte d'Ivoire") # You add the others here
cc$NAME[match(mappings, cc$NAME)] <- names(mappings)

wm <- left_join(wm, cc[,c("NAME","MYCONTINENTS", "sales")], by=c("region"="NAME"))
p <- ggplot() +
  geom_polygon(aes(x=long, y=lat, group=group, fill=sales), wm, colour = NA) +
  coord_quickmap() +
  scale_fill_gradient(low="light blue", high="red") +
  theme_void() +
  labs(title="Fig 4.global sales map by platform and year")
library(plotly)
# ggplotly(p)
p


# ## Download data from Natural Earth
# url <- "https://www.naturalearthdata.com/http//www.naturalearthdata.com/download/50m/cultural/ne_50m_admin_0_countries.zip"
# 
# tmp <- tempdir()
# 
# file <- basename(url)
# 
# download.file(url, file)
# 
# unzip(file, exdir = tmp)
# 
# library(rgdal)
# ## Read the data into R
# countries <- readOGR(dsn=tmp,
#                      layer = "ne_50m_admin_0_countries", 
#                      encoding = "UTF-8")
# 
# 
# t_2017<-sp::merge(countries,cc,
#                   by.y="ISO3",by.x="ISO_A3",sort=FALSE,duplicateGeoms =
#                   TRUE,all.x=FALSE)
# 
# factpal <- colorFactor(topo.colors(4), t_2017$sales)
# 
# 
# library(leaflet)
# leaflet() %>%
#   addTiles() %>%
#   addPolygons(stroke = FALSE,data=t_2017,
#   fillColor = ~factpal(t_2017$sales),
#   smoothFactor = 0.2,
#   weight = 2,
#   opacity = 1,
#   color = "white",
#   dashArray = "3",
#   fillOpacity = 0.5) 
```

Fig 4 is a global heat map to visualize sales revenue by platform and year. I take Platform="3DS" and Year_of_Release="2016" as example. As we can see, Japan had highest sales revenue for 3DS video games in 2016. Users could also adjust Platform and Year of Release in Shiny App. 

##5.game name text analysis (word cloud)
```{r,echo=FALSE,message=FALSE,warning=FALSE}
##Take Year_of_Release="2016" as example
library(tidytext)
library(dplyr)
text <- mydata %>%
  filter(Year_of_Release == 2016) %>%
  select(Name) %>%
  unnest_tokens(word, Name)

data(stop_words)
tidy_text <- text %>%
  anti_join(stop_words)

library(tm)
text <- toString(tidy_text)
docs <- Corpus(VectorSource(text))
toSpace <- content_transformer(function (x , pattern ) gsub(pattern, " ", x))
docs <- tm_map(docs, toSpace, "/")
docs <- tm_map(docs, toSpace, "@")
docs <- tm_map(docs, toSpace, "\\|")
docs <- tm_map(docs, removeNumbers)
# Remove punctuations
docs <- tm_map(docs, removePunctuation)
# Eliminate extra white spaces
docs <- tm_map(docs, stripWhitespace)
dtm <- TermDocumentMatrix(docs)
m <- as.matrix(dtm)
v <- sort(rowSums(m),decreasing=TRUE)
d <- data.frame(word = names(v),freq=v)
# head(d, 10)

library(wordcloud)
layout(matrix(c(1, 2), nrow=2), heights=c(1, 4))
par(mar=rep(0, 4))
plot.new()
text(x=0.5, y=0.5, "Fig 5. Word Cloud")
set.seed(321)
wordcloud(words = d$word, freq = d$freq, min.freq = 1,
          max.words=100, random.order=FALSE, rot.per=0.35,
          colors=brewer.pal(8, "Dark2"))


# layout(matrix(c(1, 2), nrow=2), heights=c(1, 4))
# par(mar=rep(0, 4))
# plot.new()
# text(x=0.5, y=0.5, "Fig 5. Word Cloud")
# tidy_text%>%
#   count(word) %>%
#   with(wordcloud(word, n, min.freq = 1,max.words = 100,random.order=FALSE, 
#                  rot.per=0.35,main="Title",colors=brewer.pal(8, "Dark2")))
```

Fig 5 is a word cloud allows us to highlight the most frequently used keywords in texts. The above word cloud clearly shows that “collection”, “lego”, “wars”, “pro” and “evil” are the five most important words in game names which released 2016.

##6.Sales by Genre
```{r,echo=FALSE}
##Take Genre="Action" as an example
mydata$Genre <- as.character(mydata$Genre)
games_y1<-mydata %>%
  filter(Genre == "Action") %>%
  select(Year_of_Release,EU_Sales,JP_Sales,NA_Sales,Other_Sales)

games_long<-gather(games_y1,Region,TotalSales,EU_Sales:Other_Sales)

q <- ggplot(games_long,aes(x=Year_of_Release,y=TotalSales,fill=Region))+
  geom_bar(stat="identity")+
  labs(title="Fig 6.Sales by Genre")
ggplotly(q)
```

Fig 6 shows the distribution of sales by genre. It is clear that North America always contributes most of sales, however in 2015 and 2016, it seems all four regions had equal sales amount.

##7.Pie chart of genres in certain year and platform
```{r,echo=FALSE}
##Take Platform="3DS" and Year_of_Release="2016" as example
genre_count <- mydata %>%
  select(Year_of_Release,Platform,Genre) %>%
  filter(Year_of_Release==2016 & Platform=="3DS") %>%
  group_by(Genre) %>%
  summarise(count=n())

genre_pie <- plot_ly(genre_count, labels = ~Genre, values = ~count, type = 'pie',
                     textposition = 'inside',textinfo = 'percent+label') %>%
  layout(title="Fig 7.Pie chart of genres")

ggplotly(genre_pie)
```

Fig 7 is pie chart of genres in certain year and platform. I took Platform="3DS" and Year_of_Release="2016" as example, it shows us Action is the largest portion. Same as the global sales map above(Fig 1), Platform and Year of Release are also adjustable in Shiny App. 


#Top 10 Games
```{r}
#Take 2016 as example
toptable <- mydata %>%
  select(Name,Global_Sales,Year_of_Release,Platform) %>%
  filter(Year_of_Release==2016) %>%
  arrange(desc(Global_Sales))%>%
  select(Name,Platform,Global_Sales)%>%
  head(10)
toptable
```

I also made a table to show users that most popular games in a certain year, here is an example when "2016" is selected. 

#Benford Law
```{r}
library(benford.analysis)
bfd <- benford(mydata$Global_Sales)
plot(bfd)

library(BenfordTests)
# Euclidean Distance Test for Benford’s Law
edist.benftest(mydata$Global_Sales)
```

The p-value is smaller than 0.05 so that we reject the null hypothesis. Therefore, the goodness-of-fit test based on the Euclidean distance between the first digits' distribution and Benford's distribution shows the data does not conform to Benford's law very well.











